---
title: "DIG Portfolio Opt"
author: "Mark Metrovich"
date: "2023-04-05"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this report I will go through the use of Modern Portfolio Theory to optimize a portfolio of any number of stock holdings by measuring the annual returns against the annualized risk, while recording the Sharpe Ratio of each portfolio.
```{r, include = FALSE}

# first load packages tidyquant, plotly, magrittr, xts, zoo, timetk, tidyr, sys
# quantmod, and any other necessary ones

library(tidyquant)
library(magrittr)
library(xts)
library(plotly)
library(quantmod)
library(tidyr)
library(timetk)
library(ggplot2)
library(forcats)
```
Select tickers that make up your portfolio and assign the price data.


```{r}
tick <- c('RSP', 'V', 'JNJ', 'DIS')
#tick <- c('DIS', 'BRK-B', 'SCHW', 'IBM', 'AMZN')
price_data <- tq_get(tick, from = '2019-02-23',
                     get = 'stock.prices')
## change the from date if you want further historical data, I like to do 
## around 10 years

## enter the weights of your portfolio make sure the weights correspond to the 
## order of tickers entered in tick above 
#my_wts <- c(0.03965190376,0,0,0,0,0,0,0,0, 0.017817,0.025436,0,0.05172402,0,0.027469848,
            #0,0,0.02137,0.0314703,0.058565,0.036907,0.0239947,0.0241484,0.10025,0.4194799,0.083148,
            #0.0182638,0.020297)
my_wts <- c(0.15, 0.15, 0.4, 0.3)


```

Next transform the data using log scale
```{r}
log_ret_tidy <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = 'daily',
               col_rename = 'ret',
               type = 'log')

head(log_ret_tidy)


log_ret_xts <- log_ret_tidy %>%
  spread(symbol, value = ret) %>%
  tk_xts()

head(log_ret_xts)
tail(log_ret_xts)
```


Next calculate the mean returns, and covariance matrix

```{r}
mean_ret <- colMeans(log_ret_xts, na.rm = TRUE)
print(round(mean_ret, 5))


cov_mat <- cov(log_ret_xts, use= 'complete') * 252

print(round(cov_mat,4))
```


Now we will find some random weights for the holdings in the portfolio, as well as finding that weighted portfolio's returns, risk, and Sharpe ratio.

```{r}
wts <- runif(n = length(tick))

wts <- wts/sum(wts)
print(wts)
sum(wts)

port_returns <- (sum(wts * mean_ret) + 1)^252 - 1
port_returns # we see an ___ return


port_risk <- sqrt(t(wts) %*% (cov_mat %*% wts))
print(port_risk) # we see a ____ 

# Since Risk free rate is 0% 

sharpe_ratio <- port_returns/port_risk
print(sharpe_ratio) # sharpe ratio is ____



```


Once we have made sure the work is correct, we begin the function to run the 5000 (can use different number) trials. Firstly, create the blank vectors within the DF to store numbers.

```{r}
num_port <- 15000

# Creating a matrix to store the weights

all_wts <- matrix(nrow = num_port,
                  ncol = length(tick))

# Creating an empty vector to store
# Portfolio returns

port_returns <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Standard deviation

port_risk <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Sharpe Ratio

sharpe_ratio <- vector('numeric', length = num_port)


```


Now we begin the loop 

```{r}
for (i in seq_along(port_returns)) {
  
  wts <- runif(length(tick))
  wts <- wts/sum(wts)
  
  # Storing weight in the matrix
  all_wts[i,] <- wts
  
  # Portfolio returns
  
  port_ret <- sum(wts * mean_ret)
  port_ret <- ((port_ret + 1)^252) - 1
  
  # Storing Portfolio Returns values
  port_returns[i] <- port_ret
  
  
  # Creating and storing portfolio risk
  port_sd <- sqrt(t(wts) %*% (cov_mat  %*% wts))
  port_risk[i] <- port_sd
  
  # Creating and storing Portfolio Sharpe Ratios
  # Assuming 0% Risk free rate
  
  sr <- port_ret/port_sd
  sharpe_ratio[i] <- sr
  
}
```


Now we store the values in a table 

```{r}
# Storing the values in the table
portfolio_values <- tibble(Return = port_returns,
                           Risk = port_risk,
                           SharpeRatio = sharpe_ratio)


# Converting matrix to a tibble and changing column names
all_wts <- tk_tbl(all_wts)

colnames(all_wts) <- colnames(log_ret_xts)

# Combing all the values together
portfolio_values <- tk_tbl(cbind(all_wts, portfolio_values))

head(portfolio_values)
```

Now we will utilize our weights inputted at the top and view their statistics. 

```{r}
my_wts <- my_wts/sum(my_wts)
print(my_wts)
sum(my_wts)

my_port_returns <- (sum(my_wts * mean_ret) + 1)^252 - 1
my_port_returns # we see our return 

my_port_risk <- sqrt(t(my_wts) %*% (cov_mat %*% my_wts))
print(my_port_risk) # we see our risk

# Since Risk free rate is 0% 

my_sharpe_ratio <- (my_port_returns-0.0525)/my_port_risk
print(my_sharpe_ratio) # sharpe ratio 

my_info <- c(my_wts, my_port_returns, my_port_risk, my_sharpe_ratio)
my_labels <- c(tick, "Return", "Risk", "SharpeRatio")
my_port <- data.frame(my_labels, my_info)


```

Then find the portfolios with min variance, optimal Sharpe ratio, and max returns

```{r}
min_var <- portfolio_values[which.min(portfolio_values$Risk),]
max_sr <- portfolio_values[which.max(portfolio_values$SharpeRatio),]
max_ret <- portfolio_values[which.max(portfolio_values$Return),]
```


Then can create some graphs and charts to see our data


Histograms that reflect holdings in the three selected portfolios
```{r}

# min variance
p <- min_var %>%
  gather(tick, key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Minimum Variance Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 

ggplotly(p)

### max returns
p <- max_ret %>%
  gather(tick, key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Maximum Returns Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 

ggplotly(p)


### optmizied sharpe ratio
p <- max_sr %>%
  gather(tick, key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Tangency Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 

ggplotly(p)
```

Then we can plot the efficient frontier and view how the portfolios fair on a complete scale

```{r}
p <- portfolio_values %>%
  ggplot(aes(x = Risk, y = Return, color = SharpeRatio)) +
  geom_point() +
  theme_classic() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = 'Annualized Risk',
       y = 'Annualized Returns',
       title = "Portfolio Optimization & Efficient Frontier") +
  geom_point(aes(x = Risk,
                 y = Return), data = min_var, color = 'red') +
  geom_point(aes(x = Risk,
                 y = Return), data = max_sr, color = 'red')+
  annotate('text', x = max_sr$Risk, y = max_sr$Return, label = "Tangency Portfolio") +
  annotate('text', x = min_var$Risk + 0.025, y = min_var$Return, label = "Minimum variance portfolio") + 
  geom_point(aes(x = my_port_risk, y = my_port_returns), color = 'orange') +
  annotate('text', x = my_port_risk, y = my_port_returns, label = "My Portfolio")
 # once you find your port's risk and returns, can plot labels and point on graph 
 
ggplotly(p)
```

Thus we find the efficient frontier graph, be aware that this is only using past data and there is no prediction in this code. 



------------------------------------------------------------------------------

## Portfolio Analytics 

Now that we have graphed the efficient frontier and seen how our portfolio compares to the optimized ones from 5,000 random trials. Let's look at where we stand numerically, for some simplicity if we do not want to have to look at our random plots for each and every case. 

We will look at the portfolio returns annualized, as well as it's calculated risk, sharpe ratio, and how it compares to the benchmark. This may be more suitable for situations where we want to see how a single added holding influences these factors based on historical data. 

Lets load our packages first 

```{r}
library(PortfolioAnalytics)
library(fPortfolio)
```


```{r}
# lets bind our portfolio 
portfolioPrices <- NULL
for (ticker in tick) {
  portfolioPrices <- cbind(portfolioPrices,
                           getSymbols.yahoo(ticker,
                                            from="2023-07-25", 
    periodicity = 'daily', auto.assign = FALSE)[,4])
  
  # us taking the close data
}

# check if missing data 

colSums(is.na(portfolioPrices)) # checking how many NA values over the time period


```
Now we have our prices over the desired period, lets go ahead and find the returns. This gives us the portfolio returns each day as the ROC() function takes the rate of change at each close.


```{r}
portfolioRets <- na.omit(ROC(portfolioPrices)) # want to omit the NA values
```

For our comparison to the benchmark we will be referring to the S&P 500, by using the SPY ticker. 

```{r}
benchmarkPrices <- getSymbols.yahoo('SPY', from = '2023-07-25',
                                    periodicity = 'daily',
                                    auto.assign = FALSE)
benchmarkRet <- na.omit(dailyReturn(benchmarkPrices$SPY.Close, type = "log"))
```

Now we have portfolio returns and the benchmark returns, now we want to aggregate the returns using our weights to show the actual port returns

```{r}
portfolioReturn <- Return.portfolio(portfolioRets)
# this is using geometric returns, base is TRUE
```

Now that we have all this stuff, we can put together some of our metrics for the portfolio like beta and alpha. This is all pretty much using the performance analytics


```{r}
print(CAPM.beta(portfolioReturn, benchmarkRet, .035/252))
# the risk free rate in this case is some number over 252
# 252 represents the number of trading days in a yr

print(CAPM.jensenAlpha(portfolioReturn, benchmarkRet, 0.035/252))

SharpeRatio(portfolioReturn, 0.035/252)
# we want the STDDev Sharpe

print(SharpeRatio(portfolioReturn, 0.035/252)[1,])

```


Now we can show some returns annualized and calendar year


```{r}
table.AnnualizedReturns(portfolioReturn)
table.CalendarReturns(portfolioReturn)

```



