---
title: "Portfolio Optimization"
author: "Mark Metrovich"
date: "2023-04-05"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this report I will go through the use of Modern Portfolio Theory to optimize a portfolio of any number of stock holdings by measuring the annual returns against the annualized risk, while recording the Sharpe Ratio of each portfolio.
```{r, include = FALSE}

# first load packages tidyquant, plotly, magrittr, xts, zoo, timetk, tidyr, sys
# quantmod, and any other necessary ones

library(tidyquant)
library(magrittr)
library(xts)
library(plotly)
library(quantmod)
library(tidyr)
library(timetk)
library(ggplot2)
library(forcats)
```
Select tickers that make up your portfolio and assign the price data.


```{r}
tick <- c('AMZN', 'BBJP', 'BLD', 'FCX', 'FLUT',
          'GOOGL', 'IHI', 'MSFT', 'NLR', 'PSA', 
          'RSP', 'SCHW', 'SNOW', 'TSM', 'VDIGX',
          'VOO', 'VTV')
price_data <- tq_get(tick, from = '2014-01-01',
                     get = 'stock.prices')
## change the from date if you want further historical data, I like to do 
## around 10 years

## enter the weights of your portfolio make sure the weights correspond to the 
## order of tickers entered in tick above 
my_wts <- c(3.58, 2.48, 2.06, 2.68, 1.8, 
            5.44, 3.39, 2.36, 1.59, 1.65,
            2.74, 1.87, 2.22, 5.30, 6.88,
            48.05, 5.85
            )/100
```

Next transform the data using log scale
```{r}
log_ret_tidy <- price_data %>%
  group_by(symbol) %>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period = 'daily',
               col_rename = 'ret',
               type = 'log')

head(log_ret_tidy)


log_ret_xts <- log_ret_tidy %>%
  spread(symbol, value = ret) %>%
  tk_xts()

head(log_ret_xts)
tail(log_ret_xts)
```


Next calculate the mean returns, and covariance matrix

```{r}
mean_ret <- colMeans(log_ret_xts, na.rm = TRUE)
print(round(mean_ret, 5))


cov_mat <- cov(log_ret_xts, use= 'complete') * 252

print(round(cov_mat,4))
```


Now we will find some random weights for the holdings in the portfolio, as well as finding that weighted portfolio's returns, risk, and Sharpe ratio.

```{r}
wts <- runif(n = length(tick))

wts <- wts/sum(wts)
print(wts)
sum(wts)

port_returns <- (sum(wts * mean_ret) + 1)^252 - 1
port_returns # we see an ___ return


port_risk <- sqrt(t(wts) %*% (cov_mat %*% wts))
print(port_risk) # we see a ____ 

# Since Risk free rate is 0% 

sharpe_ratio <- port_returns/port_risk
print(sharpe_ratio) # sharpe ratio is ____



```


Once we have made sure the work is correct, we begin the function to run the 5000 (can use different number) trials. Firstly, create the blank vectors within the DF to store numbers.

```{r}
num_port <- 15000

# Creating a matrix to store the weights

all_wts <- matrix(nrow = num_port,
                  ncol = length(tick))

# Creating an empty vector to store
# Portfolio returns

port_returns <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Standard deviation

port_risk <- vector('numeric', length = num_port)

# Creating an empty vector to store
# Portfolio Sharpe Ratio

sharpe_ratio <- vector('numeric', length = num_port)


```


Now we begin the loop 

```{r}
for (i in seq_along(port_returns)) {
  
  wts <- runif(length(tick))
  wts <- wts/sum(wts)
  
  # Storing weight in the matrix
  all_wts[i,] <- wts
  
  # Portfolio returns
  
  port_ret <- sum(wts * mean_ret)
  port_ret <- ((port_ret + 1)^252) - 1
  
  # Storing Portfolio Returns values
  port_returns[i] <- port_ret
  
  
  # Creating and storing portfolio risk
  port_sd <- sqrt(t(wts) %*% (cov_mat  %*% wts))
  port_risk[i] <- port_sd
  
  # Creating and storing Portfolio Sharpe Ratios
  # Assuming 0% Risk free rate
  
  sr <- port_ret/port_sd
  sharpe_ratio[i] <- sr
  
}
```


Now we store the values in a table 

```{r}
# Storing the values in the table
portfolio_values <- tibble(Return = port_returns,
                           Risk = port_risk,
                           SharpeRatio = sharpe_ratio)


# Converting matrix to a tibble and changing column names
all_wts <- tk_tbl(all_wts)

colnames(all_wts) <- colnames(log_ret_xts)

# Combing all the values together
portfolio_values <- tk_tbl(cbind(all_wts, portfolio_values))

head(portfolio_values)
```

Now we will utilize our weights inputted at the top and view their statistics. 

```{r}
my_wts <- my_wts/sum(my_wts)
print(my_wts)
sum(my_wts)

my_port_returns <- (sum(my_wts * mean_ret) + 1)^252 - 1
my_port_returns # we see our return 

my_port_risk <- sqrt(t(my_wts) %*% (cov_mat %*% my_wts))
print(my_port_risk) # we see our risk

# Since Risk free rate is 0% 

my_sharpe_ratio <- (my_port_returns-0.0525)/my_port_risk
print(my_sharpe_ratio) # sharpe ratio 

my_info <- c(my_wts, my_port_returns, my_port_risk, my_sharpe_ratio)
my_labels <- c(tick, "Return", "Risk", "SharpeRatio")
my_port <- data.frame(my_labels, my_info)


```

Then find the portfolios with min variance, optimal Sharpe ratio, and max returns

```{r}
min_var <- portfolio_values[which.min(portfolio_values$Risk),]
max_sr <- portfolio_values[which.max(portfolio_values$SharpeRatio),]
max_ret <- portfolio_values[which.max(portfolio_values$Return),]
```


Then can create some graphs and charts to see our data


Histograms that reflect holdings in the three selected portfolios
```{r}

# min variance
p <- min_var %>%
  gather(tick, key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Minimum Variance Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 

p

### max returns
p <- max_ret %>%
  gather(tick, key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Maximum Returns Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 

p


### optmizied sharpe ratio
p <- max_sr %>%
  gather(tick, key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Tangency Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 

p
```

Then we can plot the efficient frontier and view how the portfolios fair on a complete scale

```{r}
p <- portfolio_values %>%
  ggplot(aes(x = Risk, y = Return, color = SharpeRatio)) +
  geom_point() +
  theme_classic() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = 'Annualized Risk',
       y = 'Annualized Returns',
       title = "Portfolio Optimization & Efficient Frontier") +
  geom_point(aes(x = Risk,
                 y = Return), data = min_var, color = 'red') +
  geom_point(aes(x = Risk,
                 y = Return), data = max_sr, color = 'red')+
  annotate('text', x = max_sr$Risk, y = max_sr$Return, label = "Tangency Portfolio") +
  annotate('text', x = min_var$Risk * 1.05, y = min_var$Return, label = "MIN VAR P") + 
  geom_point(aes(x = my_port_risk, y = my_port_returns), color = 'orange') +
  annotate('text', x = my_port_risk, y = my_port_returns, label = "My Portfolio")
 # once you find your port's risk and returns, can plot labels and point on graph 
p

```

Thus we find the efficient frontier graph, be aware that this is only using past data and there is no prediction in this code. 
